---
title: "PA1_template"
author: "hsjeon"
date: "Sunday, April 19, 2015"
output: html_document
---

#Assignment

##Loading and preprocessing the data

###1. Load the data
###2. Process/transform the data(if necessary) into a forat suitable for your analysis

```{r, echo=TRUE}
setwd("E:/")
activity <- read.csv("E:/activity.csv", header=TRUE)
head(activity)
```

##What is mean total number of steps taken per day?

###1. Calculate the total number of steps taken per day

```{r, echo=TRUE}
t.steps.per.day <- aggregate(activity$steps, by=list(DATE=activity$date), FUN=sum)
head(t.steps.per.day)
```

###2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
hist(t.steps.per.day[,2], col="lightpink", border="lightgrey",
     main="Histogram of the Total Number of Steps Taken Per Day", 
     xlab="Total number of steps taken per day", ylab="Frequency")
```

###3. Calculate and report the mean and median of the total number of steps taken per day

```{r, echo=TRUE}
summary(t.steps.per.day[,2])
```

##What is the average daily activity pattern?

###1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
average.steps <- aggregate(activity$steps, by=list(INTERVAL=activity$interval), 
			   FUN=mean, na.rm=TRUE)
head(average.steps)
        
plot(average.steps, type="l", main="Average Daily Activity Pattern",
     xlab="5-minute interval", ylab="Average number of steps")
```

###2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}
subset(average.steps, grepl(max(average.steps[,2]), average.steps$x))
```

##Imputing missing values
###1. Calculate and report the total number of missing values in the dataset

```{r, echo=TRUE}
length(activity$steps[activity$steps=="NA"]) 
```

###2. Devise a strategy for filling in all of the missing values in the dataset

```{r, echo=TRUE}
activity$steps[is.na(activity$steps)] <- mean(average.steps[,2], na.rm=TRUE)        
```

###3. Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r, echo=TRUE}
impute.activity <- activity
head(impute.activity)
```

###4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
impute.t.steps.per.day <- aggregate(impute.activity$steps, 
                                            by=list(DATE=impute.activity$date), FUN=sum)
head(impute.t.steps.per.day) 

hist(impute.t.steps.per.day[,2], col="lightblue", border="lightgrey",
     main="New Histogram of the Total Number of Steps Taken Per Day", 
     xlab="New total number of steps taken per day", ylab="Frequency")

summary(impute.t.steps.per.day[,2])
```

##Are there differences in activity patterns between weekdays and weekends?

###1. Create a new factor variable in the dataset with two levels-"weekday" and "weekend" indicating whether a given date is a weekday or weekwnd day

```{r, echo=TRUE}
Sys.setlocale("LC_TIME", "English")

impute.activity$date <- as.POSIXlt(impute.activity$date, format="%Y-%m-%d")
impute.activity$day <- weekdays(impute.activity$date) 

day.impute.activity <- impute.activity
head(day.impute.activity)
        
day.impute.activity$day <- gsub("(Monday)|(Tuesday)|(Wednesday)|(Thursday)|(Friday)", 
				"weekday", impute.activity$day)
day.impute.activity$day <- gsub("(Saturday)|(Sunday)", "weekend", day.impute.activity$day) 

week.impute.activity <- day.impute.activity
head(week.impute.activity)
        
weekday.impute.activity <- subset(week.impute.activity, week.impute.activity$day==c("weekday"))
head(weekday.impute.activity)

weekend.impute.activity <- subset(week.impute.activity, week.impute.activity$day==c("weekend"))
head(weekend.impute.activity)

weekday.average.steps <- aggregate(weekday.impute.activity$steps, 
				   by=list(INTERVAL=weekday.impute.activity$interval), FUN=mean)
head(weekday.average.steps)

weekend.average.steps <- aggregate(weekend.impute.activity$steps, 
				   by=list(INTERVAL=weekend.impute.activity$interval), FUN=mean)
head(weekend.average.steps)
```

###2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
   
```{r, echo=TRUE}
par(mfrow = c(2, 1))
       
plot(weekday.average.steps, type="l", col="black", main="Average Weekday Activity Pattern",
     xlab="5-minute interval", ylab="Average number of steps")
plot(weekend.average.steps, type="l", col="red", main="Average Weekend Activity Pattern",
     xlab="5-minute interval", ylab="Average number of steps")  
```